---
title: "Coding Scheme with Perspective, Message Type and Not-serious/Unclear Dimensions: Tweet Frequencies"
author: "Hannah Metzler"
date: "12/05/2021"
output: 
  pdf_document:
    df_print: kable
    keep_tex: true
url_colour: blue
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
Sys.setlocale("LC_ALL", 'en_US.UTF-8')
options(scipen=99)
library(dplyr)
library(stringr)
library(irr)
library(xfun)
library(caret)
library(forcats)
library(RColorBrewer)
```

# Add 1000 randomly selected tweets to hand- and model-selected training set. Labels for focus/perspective and type of message

```{r}
training <- read.csv('../tweet_training_set/training_posts20201102_2185tweets.tsv', sep='\t') %>%  #training sample according to table with 6 content types and 2 perspectives
  select(-txt_clean)
rs <- read.csv('../tweet_training_set/1-1000tweets_basefrequency_categories_labelled.tsv', sep='\t') %>% #random sample of tweets coded according to table with content and perspective classification
  select(-Order)
#names(training); names(rs)
totaltrain <- rbind(training, rs)
```

How many tweets per category (all small categories included) in all coded tweets? 
```{r}
as.data.frame(xtabs(~Category, data=totaltrain))
#write.table(as.data.frame(xtabs(~Category, data=totaltrain)),'../output/tweets_per_category_trainingposts20201103.tsv', sep='\t', row.names=F)
```

- It looks like Coping1/3 and Suicidality1/3 will have to be combined, respectively. 
- Life saved not enough for Machine learning. 


Do all categories have the right focus coded?

- Suicide_other can be all 3 perspectives/foci
- Off-topic are all neither of them 
- Werther can rarely be about a suicide case and still focused on prevention, see 7 instances below

```{r}
as.data.frame(xtabs(~focus+Category, data=totaltrain))
# #look at the example tweets
# totaltrain %>% 
#   filter(Category=="Werther" & focus == "2") %>% 
#   select(Category, focus, Contents)
```

Which categories can be labelled as non-serious? Only suicide other or off-topic tweets. 

```{r}
as.data.frame(xtabs(~notserious_unclear+Category, data=totaltrain))
```


# Frequency of categories in a random sample of 1000 tweets

How many tweets per category (all small categories included)? 

```{r}
prop.cat = as.data.frame(xtabs(~Category, data=rs))

#define the order for factor levels
ord_category <- c( "Suicidality1","Suicidality3","Coping1", "Coping3", "News_coping", "News_suicidality", "Bereaved_negative", "Bereaved_coping","Awareness", "Prevention",  "Werther", "Life_saved", "Suicide_other","Off-topic")

coded = rs %>% 
  mutate(Category = factor(Category, levels = ord_category, labels = ord_category)) %>% 
  #combine all coping and suicidality categories into one
  mutate(Combined = recode(Category, Coping1 = "Coping", Coping3 = "Coping", Suicidality1 = "Suicidality", Suicidality3 = "Suicidality", 
                           News_coping = "Coping", News_suicidality = "Suicidality"))
```

Total number of tweets n=`r sum(prop.cat$Freq,2)`. 

Proportion of tweets per small category
```{r}
#proportion per category
prop.cat <- as.data.frame(coded %>% 
  group_by(Category) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n))) %>%
  ungroup()  %>% 
  mutate(Combined = recode(Category, Coping1 = "Coping", Coping3 = "Coping", Suicidality1 = "Suicidality", Suicidality3 = "Suicidality", 
                           News_coping = "Coping", News_suicidality = "Suicidality"))
prop.cat
```

Tweets per category if 1st and 3rd person perspective are combined for Coping and Suicidality, and news coping and news suicidality classes included as well. Total number of tweets n=`r sum(prop.cat$n,2)`.  

```{r}
#proportion with combined categories
prop.comb <- as.data.frame(coded %>%
  group_by(Combined) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n))) %>% 
  ungroup()
#write.table(prop.cat,'../output/tweets_per_category_basefrequency20201028.tsv', sep='\t', row.names=F)
prop.comb
```


```{r, Figure Frequency in random sample combined categories}
s=15
plot.prop.comb <- ggplot(prop.comb, aes(x=Combined, y = freq, fill=Combined)) + #dataset and variables to plot
  geom_bar(stat="identity")+
  ggtitle("Frequency in random tweet sample")+
  labs(y="Proportion", x="") + #axes and title labels
    scale_fill_brewer(palette="Paired")+
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position="none", text=element_text(size=s))

plot.prop.comb
#pdf('../figures/proportion_categories_random_sample.pdf', height=5, width=7); plot.prop.comb; plot.prop.coping.suicidality; dev.off()
```

Tweets in the suicidality and coping categories include the following subcategories:

```{r, Figure Frequency in random sample subcategories}

#Subcategories
df <- filter(prop.cat, Combined=="Coping" | Combined=="Suicidality")
plot.prop.coping.suicidality <- ggplot(df, aes(x=Category, y = freq, fill=Combined)) + #dataset and variables to plot
  geom_bar(stat="identity")+
  ggtitle("Subcategories Coping and Suicidality") +
  labs(y="Proportion", x="") + #axes and title labels
  scale_fill_brewer(palette="Paired")+
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position="none", text=element_text(size=s))
plot.prop.coping.suicidality



#pdf('../figures/proportion_subcategories_copingsuicidal_random.pdf', height=5, width=7/10*6); plot.prop.coping.suicidality; dev.off()
 
```
```{r}
prop.notserious <- as.data.frame(coded %>%
  group_by(notserious_unclear) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n))) %>% 
  ungroup()
```
Out of the off-topic and suicide_other tweets, about `r round(prop.notserious$freq[2], 2)*100`% are jokes, sarcastic, metaphors or 

# Frequency in training sample

```{r}
coded = totaltrain %>% 
  filter(!Category=="") %>% #filter empty (not labeled) tweets 
  mutate(Category = factor(Category, levels = ord_category, labels = ord_category)) %>% 
  mutate(Combined = recode(Category, Coping1 = "Coping", Coping3 = "Coping", Suicidality1 = "Suicidality", Suicidality3 = "Suicidality", 
                           News_coping = "Coping", News_suicidality = "Suicidality", Life_saved= "Suicide_other", Bereaved_negative = "Suicide_other", Bereaved_coping = "Suicide_other")) %>% 
  mutate(Contents = as.character(Contents))

# proportion per category
prop.cat <- as.data.frame(coded %>% 
  group_by(Category) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n))) %>%
  ungroup()  %>% 
  mutate(Combined = recode(Category, Coping1 = "Coping", Coping3 = "Coping", Suicidality1 = "Suicidality", Suicidality3 = "Suicidality", 
                           News_coping = "Coping", News_suicidality = "Suicidality"))

#proportion with combined categories
prop.comb <- as.data.frame(coded %>%
  group_by(Combined) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n))) %>% 
  ungroup()
```

Tweets per category. Total number of tweets n=`r sum(prop.comb$n,2)`. 

```{r}
prop.cat
```


Tweets per category if 1st and 3rd person perspective are combined for Coping and Suicidality. Total number of tweets n=`r sum(prop.comb$n,2)`.  
```{r}
prop.comb
```

```{r, Figure Frequency in training sample}
s=15
plot.n.comb <- ggplot(prop.cat, aes(x=Combined, y = n, fill=Combined)) + #dataset and variables to plot
  geom_bar(stat="identity")+
  ggtitle(paste("Frequency in training sample with n=", nrow(totaltrain), sep=""))+
  labs(y="Number of tweets", x="Category") + #axes and title labels
  scale_fill_brewer(palette="Paired")+
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position="none", text=element_text(size=s))

plot.n.comb

s=14
#Subcategories
df <- filter(prop.cat, Combined=="Coping" | Combined=="Suicidality")
plot.n.coping.suicidality <- ggplot(df, aes(x=Category, y = n, fill=Combined)) + #dataset and variables to plot
  geom_bar(stat="identity")+
  ggtitle(paste("Subcategories Coping & Suicidality in training sample with n=", nrow(totaltrain), sep="")) +
  labs(y="Number of tweets", x="Category") + #axes and title labels
  scale_fill_brewer(palette="Paired")+
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position="none", text=element_text(size=s))
plot.n.coping.suicidality


#pdf('../figures/n_per_category_training_sample.pdf', height=5, width=7); plot.n.comb; plot.n.coping.suicidality; dev.off()
#pdf('../figures/n_per_subcategory_copingsuicidal_training.pdf', height=5, width=7/10*6); plot.n.coping.suicidality; dev.off()
 
```

Which categories in the training sample have less than 150 tweets? We need to combined these into larger categories, or code them as Suicide_other for training the models. 

```{r, Figure small categories}

s=15
plot.smalln <- ggplot(filter(prop.cat, n < 150), aes(x=Category, y = n, fill=Combined)) + #dataset and variables to plot
  geom_bar(stat="identity")+
  ggtitle("Small categories training sample")+
  labs(y="Number of tweets", x="Category") + #axes and title labels
  scale_fill_brewer(palette="Paired")+
  theme_bw() +  
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position="none", text=element_text(size=s))

plot.smalln

#pdf('../figures/small_categories_training_sample.pdf', height=5, width=7); plot.smalln; dev.off()

```


# Save the complete training set: training_posts20201105_3181tweets.tsv

```{r}
# write this dataframe to new training sample file. Very few tweets have \n new line characters that split the tweet into 2 cells in the ouput.tsv file - delete these characters first
# example tweets
str_subset(coded$Contents, "It's a cry for help that went unanswered.") # "in over half of them", "It's a cry for help that went unanswered."
# problem with this: some tweets have new lines, they need to be deleted before I can write a tab seperated file, otherwise the tweet text gets split into news cells and lines at these points
coded$Contents = str_replace_all(coded$Contents, "\n","")
#write.table(coded,'../tweet_training_set/training_posts20201105_3181tweets.tsv', sep = "\t", row.names=F) 
```
  


