---
title: "Tweets per Category"
author: "Hannah Metzler"
date: "3/16/2021"
output: 
  pdf_document:
    df_print: kable
    keep_tex: true
url_colour: blue
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, results=T)
Sys.setlocale("LC_ALL", 'en_US.UTF-8')
options(scipen=99)
library(dplyr)
library(ggplot2)
library(viridis)
#load data (origin: scripts folder, where this .Rmd file is)
load("../tweets_other_datasets/Suicide_tweets_daily_volume_per_maincategory.R")
#our source script (but this takes a couple of minutes)
# source('06BuildPredictionsTimelineDataset.R')

#assign colours to categories: 
catcols = viridis(6)
mycols = c("awareness"= catcols[1], "werther"=catcols[6], "prevention"=catcols[4], "suicidality"=catcols[5], "coping"=catcols[2], "irrelevant"=catcols[3])

#delete the month year variable (does not function properly)
dcatperday$monthyear=NULL
```

# Dataset description

We initially downloaded tweets from crimson hexagon on April 15th 2020, including all tweets containing a suicide-related term from January 1st 2013 to April 14th 2020. We excluded terms that clearly do not refer to suicide as the act of killing oneself. The query on Crimson Hexagon was: 

  country:USA AND NOT engagementType:RETWEET AND language:en
  AND (suicide OR suicidal OR "killed himself" OR "killed herself" OR "kill himself" OR "kill herself" OR "hung himself" OR "hung herself" OR "took his life" OR "took her life" OR "take his life" OR "take her life" OR "end his own life" OR "end her own life" OR "ended his own life" OR "ended her own life" OR "end his life" OR "end her life" OR "ended his life" OR "ended her life" OR "ends his life" OR "ends her life") 
  AND NOT ("suicide squad" OR suicidechrist OR suicidegirl* OR suicideboy* OR suicideleopard OR suicidexjockey* OR "suicidal grind" OR bomber OR squad OR epstein OR Trump OR clinton* OR Hillary OR Biden OR sanders OR “political suicide”) 
  AND NOT ("RT" OR "MT")

We rehydrated these tweets a first time in April 2020, the resulting dataset included around 13.428 million tweets, of which only 9.584 million had unique ids. The machine learning models were trained on a random sample of manually annotated tweets from this sample. To get the number of retweets, we rehydrated a second time on March 14th 2021, which produced a sample of 8.892 million unique tweets. We then added the number of retweets to each tweet, to count the total number of times it was posted on Twitter. (Side note: the retweets are only on that date, but most retweets on Twitter happen with a couple of hours.)

Predictions were made with the BERT model (check with Hubert). 

To do: 
- make predictions for the variable about suicide vs. off-topic


# Total number of tweets per category by year

* Irrelevant increase over the years may simply reflect increase in tweets overall. Irrelevant includes tweets about suicide not belonging in any other category, and off-topic tweet (metaphors, jokes, euthansia, bombings,...). 
* increasing awareness efforts especially in 2019?
* increasing awareness and prevention efforts over the years

```{r}
dcatperyear = dcatperday %>% 
  group_by(year, predictions) %>% 
  summarise(ntweets = sum(ntweets)) %>% 
  #2020 is not a full year 
  filter(year!="2020") %>% 
  #million tweets
  mutate(thousand_tweets = ntweets/1000)

dtotperyear = dcatperday %>%
  group_by(date) %>% #only first line of 6 categories (ntotal is the same on each line)
  slice(1) %>% 
  select(-predictions) %>% 
  group_by(year) %>% 
  summarise(ntotal = sum(ntotal)) %>% 
  #2020 is not a full year 
  filter(year!="2020") %>% 
  #million tweets
  mutate(thousand_tweets = ntotal/1000)

#plot per year
ggplot(data=dcatperyear, aes(x=year, y=thousand_tweets))+
  geom_point(aes(colour = predictions))+ geom_line(aes(colour = predictions))+
  #show total tweets with suicide terms per year (sum of all categories)
  # geom_line(data=dtotperyear, aes(x=year, y=ntotal))+
  theme_bw()+
  scale_colour_manual(values=mycols)+
  scale_x_continuous(n.breaks=7)
```

 
 
```{r}
#plot per year
ggplot(data=filter(dcatperyear, predictions!="irrelevant"), aes(x=year, y=thousand_tweets, colour = predictions))+
  geom_point()+ geom_line()+
  theme_bw()+
  scale_colour_manual(values=mycols)+
  scale_x_continuous(n.breaks=7)+
  ggtitle(" Same plot without the category Irrelevant")
  
```

* also shows the general increase for Werther tweets. 
* the increase exists since 2017, but is much smaller, for coping and suicidality tweets


# Monthly pattern? Total number of tweets per month (averaged across 7 years)

```{r}
dcatpermonth = dcatperday %>% 
  filter(date < as.Date("2020-04-01")) %>% #no full month of data
  group_by(month, predictions) %>% 
  summarise(ntweets = sum(ntweets)) %>% 
  #thousand tweets
  mutate(thousand_tweets = ntweets/1000) %>% ungroup()
#plot per month
ggplot(data=dcatpermonth, aes(x=month, y=thousand_tweets, colour = predictions))+
  geom_point()+ geom_line(aes(group=predictions))+
  theme_bw()+
  scale_colour_manual(values=mycols)

```

* Open question: Are the peaks in certain categories driven by outlier months in some years, or general patterns?
    - September: world suicide prevention day is likely a general pattern
    - Werther probably influenced by celebrity suicides in July, November, December in one or some of the years.
    - Could celebrity suicides also explain the high peak of awareness in June?

# Monthly timeline across years (January 2013 to March 2020)

```{r}
#monthly timeseries
dtsmonthly = dcatperday %>% 
  #filter(date < as.Date("2020-04-01")) %>% #no full month of data
  group_by(year, month, predictions) %>% 
  summarise(ntweets = sum(ntweets)) %>% 
  #thousand tweets
  mutate(thousand_tweets = ntweets/1000) %>% ungroup() %>% 
#add median per month 
  group_by(month, predictions) %>% 
  mutate(ntweets_median = median(ntweets))

#plot per month
ggplot(data=dtsmonthly)+
  geom_point(aes(x=month, y=ntweets, colour = predictions))+ 
  geom_line(aes(x=month, y=ntweets_median, group=predictions, colour=predictions))+
  theme_bw()+
  scale_colour_manual(values=mycols)+
  ggtitle("One dot per year, median as a line")
```

* This shows that the peaks of awareness in June, Werther in July, Nov, Dec were driven by one single year (likely celebrity deaths).
* The suicidality peak in August is also driven by one single year. 

More detailed picture (without irrelevant category):

```{r}
ggplot(data=filter(dtsmonthly, predictions!="irrelevant"))+
  geom_point(aes(x=month, y=ntweets, colour = predictions))+ 
  geom_line(aes(x=month, y=ntweets_median, group=predictions, colour=predictions))+
  theme_bw()+
  scale_colour_manual(values=mycols)+
  ggtitle("One dot per year, median as a line (without irrelevant)")
```


* Werther July  peak seems driven by 2 events
* Werther November  peak seems driven by 2 events
* Werther December  peak seems driven by 1 event


```{r}
ggplot(data=filter(dtsmonthly, predictions!="irrelevant"))+
  geom_point(aes(x=month, y=ntweets, colour = predictions))+ 
  geom_line(aes(x=month, y=ntweets_median, group=predictions, colour=predictions))+
  theme_bw()+
  scale_colour_manual(values=mycols)+
  ylim(c(0,12000))+
  ggtitle("One dot per year, median as a line, below 10.000 monthly tweets")

ggplot(data=filter(dtsmonthly, predictions!="irrelevant"))+
  # geom_violin(aes(x=month, y=ntweets, colour = predictions))+ 
  geom_point(aes(x=month, y=ntweets, colour = predictions))+ 
  geom_line(aes(x=month, y=ntweets_median, group=predictions, colour=predictions))+
  facet_wrap(~predictions)+
  theme_bw()+
  scale_colour_manual(values=mycols)+
  ylim(c(0,16000))+
  theme(axis.text.x=element_text(angle=60, hjust=1), axis.title.x=element_blank(), legend.position="none")+
  ggtitle("One dot per year, median as a line, below 16.000 monthly tweets")
```

* This shows that the coping peak in May is one outlier only. 
* September prevention is systematically higher (world suicide prevention day)

# Entire timeline of daily tweets per category

```{r}
#plot per year
ggplot(data=dcatperday, aes(x=date, y=ntweets, colour = predictions))+
  geom_line()+
  theme_bw()+
  scale_colour_manual(values=mycols)
```

```{r, coping suicidality daily}
ggplot(data=filter(dcatperday, predictions=="coping" | predictions=="suicidality"), 
                   aes(x=date, y=ntweets, colour = predictions))+
    geom_line()+ facet_wrap(~predictions, ncol=1)+
  theme_bw()+theme(legend.position="none")+
  scale_colour_manual(values=mycols)+
  ggtitle("Only coping and suicidality")

ggplot(data=filter(dcatperday, predictions=="coping" | predictions=="suicidality"), 
                   aes(x=date, y=ntweets, colour = predictions))+
    geom_line()+ facet_wrap(~predictions, ncol=1)+
  theme_bw()+theme(legend.position="none")+
  scale_colour_manual(values=mycols)+
  ylim(c(0,800))+
  ggtitle("Only coping and suicidality below 800 per day")
```

* 7 peaks above 700 daily tweets:
    - June 9 2018: likely linked to suicide of Anthony Bourdain on June 8th, and Kate Spade June 5th

```{r}
dcatperday %>% 
 filter(predictions=="coping" | predictions=="suicidality") %>% 
  filter(ntweets>700) %>% 
  select(predictions, date, ntweets) %>% 
  arrange(predictions)
```



## Awareness, Prevention, Werther

```{r}
ggplot(data=filter(dcatperday, predictions=="awareness" | predictions=="prevention"  | predictions=="werther"), 
                   aes(x=date, y=ntweets, colour = predictions))+
  geom_line()+ facet_wrap(~predictions, ncol=1)+
  theme_bw()+theme(legend.position="none")+
  scale_colour_manual(values=mycols)
```


```{r}
ggplot(data=filter(dcatperday, predictions=="awareness" | predictions=="prevention"  | predictions=="werther"), 
                   aes(x=date, y=ntweets, colour = predictions))+
  geom_line()+ facet_wrap(~predictions, ncol=1)+
  theme_bw()+theme(legend.position="none")+
  ylim(c(0,10000))+
  scale_colour_manual(values=mycols)+
  ggtitle("Daily tweets below 10000")
```


* 7 large awareness peaks above 1500 daily tweets:

```{r}
dcatperday %>% 
  filter(predictions=="awareness") %>% 
  filter(ntweets>1500) %>% 
  select(date, ntweets)
```

* 14 large prevention peaks from 2016 -2019, above 1500 daily tweets:
    - every year on September 10th
    - in 2019: before Christmas
    - other regular events to still identify: 2015-11-09, 2016-11-09 - a regular event?
    - other irregular events to stil lidentify: 2016-10-01, 2017-02-23, 2018-03-19, 2018-04-24	

```{r}
dcatperday %>% 
  filter(predictions=="prevention") %>% 
  filter(ntweets>1500) %>% 
  select(date, ntweets)
```


* Five  large Werther peaks from 2016 -2019, above 8000 daily tweets:

```{r}
dcatperday %>% 
  filter(predictions=="werther") %>% 
  filter(ntweets>8000) %>% 
  select(date, ntweets)
```


# Weekly pattern? Total number of tweets per weekday (averaged across 7 years)

```{r}
dcatperwday = dcatperday %>% 
  group_by(weekday, predictions) %>% 
  summarise(median_tweets = median(ntweets)) %>% 
  #thousand tweets
  mutate(thousand_tweets = median_tweets/1000) %>% ungroup()
#plot per week day
ggplot(data=dcatperwday, aes(x=weekday, y=median_tweets, colour = predictions))+
  geom_point()+ geom_line(aes(group=predictions))+
  theme_bw()+
  scale_colour_manual(values=mycols)+
  ggtitle("Median tweets per weekday")
ggplot(data=filter(dcatperwday, predictions !="irrelevant"), aes(x=weekday, y=median_tweets, colour = predictions))+
  geom_point()+ geom_line(aes(group=predictions))+
  theme_bw()+
  scale_colour_manual(values=mycols)+
  ggtitle("Median tweets per weekday (without irrelevant category)")
ggplot(data=filter(dcatperwday, predictions =="coping" | predictions == "suicidality"), aes(x=weekday, y=median_tweets, colour = predictions))+
  geom_point()+ geom_line(aes(group=predictions))+
  theme_bw()+
  scale_colour_manual(values=mycols)+
  ggtitle("Median tweets per weekday: suicidality and coping category)")
```

* Slightly lower numbers on Sundays and Saturdays for irrelevant, prevention, awareness and Werther tweets. Tweets by organizations (e.g. media for Werther, Lifeline for prevention), or people more active on Twitter during weekdays. 
* Slighly lower numbers of coping and suicidality tweets during weekends suggest that people use Twitter less during weekends. 
