---
title: "Select coding samples suicide tweeets"
author: "Hannah Metzler"
date: "30/9/2020"
output: 
  pdf_document:
    df_print: kable
    keep_tex: true
url_colour: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
Sys.setlocale("LC_ALL", 'en_US.UTF-8')
options(scipen=99)
library(dplyr)
library(stringr)
```

```{r}
d <- read.csv("../tweet_training_set/reserve_samples_100k_from_suicide_tweets.tsv", sep="\t")
head(d)

```
Instead I did: 
<!-- # in terminal: head -n 2000 reserve_samples_100k_from_suicide_tweets.tsv | tail -n 1000 -->


If I want to replace the cleaned tweet text with text including users and URLs, I need to do: 

- get tweet IDs of all tweets for which I want uncleaned text (e.g. the 1000 base frequency sample, or all 100k tweets reserved for training: reserve_samples_100k_from_suicide_tweets.tsv)
- match with tweet IDs in from tweet_ids_set_rehydrated-time_id_text.tsv, which has the uncleaned text
- write into one file, matching the tweet ID


  
